Binocular combination in the autonomic nervous system

### 2025-02-10 manuscript.tex
Small change to add reference to thesis

Different classes of retinal photoreceptors, including cones, rods and melanopsin-containing intrinsically photosensitive retinal ganglion cells (ipRGCs), are directly involved in controlling and maintaining the size of the pupils (e.g.~McDougal and Gamlin, \protect\hyperlink{ref-McDougal2010}{2010}; Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}; Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Murray et al., \protect\hyperlink{ref-Murray2018}{2018}; Woelders et al., \protect\hyperlink{ref-Woelders2018}{2018}). Cones drive the initial rapid constriction of the pupils (Mathôt, \protect\hyperlink{ref-Mathot2018}{2018}), while the slower and longer activation of the ipRGCs maintains constriction over a prolonged period of time and regulates the post-illumination pupillary response (Markwell et al., \protect\hyperlink{ref-Markwell2010}{2010}; McDougal and Gamlin, \protect\hyperlink{ref-McDougal2010}{2010}). The ipRGCs are a recently discovered photoreceptor class (Provencio et al., \protect\hyperlink{ref-Provencio2000}{2000}) that express the photopigment melanopsin, and are involved in the regulation of the circadian rhythm (Panda et al., \protect\hyperlink{ref-Panda2002}{2002}; Ruby et al., \protect\hyperlink{ref-Ruby2002}{2002}), forming a major input to the PON (Dacey et al., \protect\hyperlink{ref-Dacey2003}{2003}). The first direct evidence of the involvement of the ipRGCs in the PLR was shown in melanopsin knockout rats by Lucas et al. (\protect\hyperlink{ref-Lucas2003}{2003}), resulting in the loss of the intrinsic photosensitivity of the cells and a reduced pupil constriction. Similar behaviour was later observed in primates and humans (Gamlin et al., \protect\hyperlink{ref-Gamlin2007}{2007}) using silent substitution (see Methods), where it was demonstrated that the PLR continues during light presentation even when cone and rod signalling is blocked, indicating the primary role of the ipRGCs in maintaining pupil constriction over a prolonged time.

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (Hubel and Wiesel, \protect\hyperlink{ref-Hubel1962}{1962}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to one eye (Campbell and Green, \protect\hyperlink{ref-Campbell1965}{1965}; Baker et al., \protect\hyperlink{ref-Baker2018}{2018}). At higher contrasts, the phenomenon of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (Baker et al., \protect\hyperlink{ref-Baker2007}{2007}; Moradi and Heeger, \protect\hyperlink{ref-Moradi2009}{2009}). This is explained by a process of interocular suppression that cancels out the additional excitatory drive caused by stimulating two eyes. Our recent work (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}) showed that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway. Here we measure the amplitude of pupil modulations in response to flickering stimuli presented as light flux, or directed towards the L-M cone, S-cone, and melanopsin pathways (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and in a dichoptic masking condition where the two eyes are stimulated at different frequencies. The results are interpreted using a contemporary model of binocular vision (Meese et al., \protect\hyperlink{ref-Meese2006}{2006}) implemented within a hierarchical Bayesian framework.

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies (e.g., Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}, \protect\hyperlink{ref-Spitschan2015}{2015}; Murray et al., \protect\hyperlink{ref-Murray2018}{2018}; Zele et al., \protect\hyperlink{ref-Zele2018}{2018}). The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}): monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean light flux. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany, Kassner et al., \protect\hyperlink{ref-Kassner2014}{2014}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

Silent substitution stimuli were prepared and calibrated for each participant with custom Python software (Martin et al., \protect\hyperlink{ref-Martin2023}{2023}) and Python scripts. Estimates of photoreceptor spectral sensitivities for each participant were constructed from the known photopigment absorbance spectra (Stockman and Sharpe, \protect\hyperlink{ref-Stockman2000}{2000}), taking account of the peak axial density of the respective photopigments, as well as lens (Pokorny et al., \protect\hyperlink{ref-Pokorny1987}{1987}; Stockman et al., \protect\hyperlink{ref-Stockman1999}{1999}; Stockman and Sharpe, \protect\hyperlink{ref-Stockman2000}{2000}) and macular pigment density (Bone et al., \protect\hyperlink{ref-Bone1988}{1988}; Stockman et al., \protect\hyperlink{ref-Stockman1999}{1999}), in accordance with the field size and age-dependant CIEPO06 observer model (CIE, \protect\hyperlink{ref-CIE2006}{2006}). The melanopic and rhodopic action spectra of the 32-year-old standard observer were taken from CIE S 026 (CIE, \protect\hyperlink{ref-CIE2018}{2018}) and then adjusted for age-related lens transmittance with a spectral correction function, in line with the standard. Macular pigment correction was not applied to the rhodopic and melanopic action spectra because rods are not present at the fovea and ipRGCs sit above the retinal pigment layer (Trieschmann et al., \protect\hyperlink{ref-Trieschmann2007}{2007}).

In the light flux experiment, the stimulus intensity was increased and decreased relative to the background, which we expected to modulate all photoreceptor classes (see Figure \ref{fig:spectraplots}e,i). In the L-M cone modulation experiment, we used silent substitution to increase the L-cone activity, and simultaneously decrease the M-cone activity, during the first half-cycle of the sine wave. In the second half-cycle the polarity of the modulation reversed (see Figure \ref{fig:spectraplots}f,j). The maximum available L-M contrast was approximately 10\%. In the S-cone modulation experiment, we increased and decreased S-cone-directed signals, whilst keeping activity in the other photoreceptors constant (see Figure \ref{fig:spectraplots}g,k). Our system allowed a maximum contrast of 45\%. Finally, in the melanopsin experiment, we modulated the activity of the melanopsin-containing intrinsically photoreceptive retinal ganglion cells, whilst keeping cone activity constant (see Figure \ref{fig:spectraplots}h,l). The maximum available melanopsin contrast was 22\%. We assume that the activity of rods was constant at the high background luminance intensity used here, and so did not attempt to silence rod activity in any condition, as this would have greatly reduced the available dynamic range. Splatter on nominally silenced photoreceptors was very small (see Figure \ref{fig:spectraplots}j-l), well below the levels that would be expected to generate measurable pupil modulations, although we can observe that the rods may not be saturated in the L-M and melanopsin-directed stimuli (Figure \ref{fig:spectraplots}j and \ref{fig:spectraplots}l) and could intrude in these two experiments. We also estimated activation of penumbral L and M cones (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}; Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Zele et al., \protect\hyperlink{ref-Zele2019}{2019}), which was minimal (\(\le 1.5%
\) contrast; less than splatter on the open-field cones) for melanopsin-directed stimuli. We note that the temporal frequency of our modulation (0.5Hz) is well below the range where penumbral cone activation can elicit visible percepts, and that such percepts fade after around 1 second (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}), and do not typically affect pupil responses (Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}).

@@

Before the start of each experiment, participants adjusted the objective planes of the lenses with the help of the experimenter until the stimulus was in focus and they perceived the two pieces of blackout material as one fused disc. Pupil responses to binocular temporal contrast modulations were examined in a factorial design that combined six ocular conditions and five temporal contrast levels: 6, 12, 24, 48 and 96\% of the available dynamic range. This design, similar to that used in our previous studies (Baker and Wade, \protect\hyperlink{ref-Baker2017}{2017}; Baker et al., \protect\hyperlink{ref-Baker2020}{2020}; Segala et al., \protect\hyperlink{ref-Segala2023}{2023}), was applied in four separate experiments, each with a different mode of photoreceptor stimulation. In the first three conditions, the discs flickered at 0.5 Hz, in either a monocular, binocular or dichoptic arrangement. In the dichoptic condition the non-target eye saw a flickering fixed contrast of 48\% of the available dynamic range. In the remaining three conditions (the cross-frequency conditions) one eye's disc flickered at 0.4 Hz, and the other eye's disc flickered at 0.5 Hz. This included monocular responses at 0.4 Hz, as well as binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 0.5 Hz, mask contrast of 48\% at 0.4 Hz in the other eye) arrangements. We counterbalanced presentation of the target stimulus across the left and right eyes.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Baker2021}{}%
Baker DH (2021) Statistical analysis of periodic data in neuroscience. Neurons, Behavior, Data analysis, and Theory 5 Available at: \url{https://doi.org/10.51628\%2F001c.27680}.

\leavevmode\hypertarget{ref-Baker2024}{}%
Baker DH, Hansford KJ, Segala FG, Morsi AY, Huxley RJ, Martin JT, Rockman M, Wade AR (2024) Binocular integration of chromatic and luminance signals. J Vis 24(12):7.

\leavevmode\hypertarget{ref-Baker2018}{}%
Baker DH, Lygo FA, Meese TS, Georgeson MA (2018) Binocular summation revisited: Beyond \(\sqrt{2}\). Psychol Bull 144:1186--1199.

\leavevmode\hypertarget{ref-Baker2007}{}%
Baker DH, Meese TS, Georgeson MA (2007) Binocular interaction: Contrast matching and contrast discrimination are predicted by the same model. Spat Vis 20:397--413.

\leavevmode\hypertarget{ref-Baker2020}{}%
Baker DH, Vilidaite G, McClarnon E, Valkova E, Bruno A, Millman RE (2020) Binaural summation of amplitude modulation involves weak interaural suppression. Sci Rep 10:3560.

\leavevmode\hypertarget{ref-Baker2017}{}%
Baker DH, Wade AR (2017) Evidence for an optimal algorithm underlying signal combination in human visual cortex. Cereb Cortex 27:254--264.

\leavevmode\hypertarget{ref-Barrionuevo2016}{}%
Barrionuevo PA, Cao D (2016) Luminance and chromatic signals interact differently with melanopsin activation to control the pupil light response. Journal of Vision 16(11):29.

\leavevmode\hypertarget{ref-Bone1988}{}%
Bone RA, Landrum JT, Fernandez L, Tarsis SL (1988) Analysis of the macular pigment by HPLC: Retinal distribution and age study. Investigative Ophthalmology \& Visual Science 29:843--849.

\leavevmode\hypertarget{ref-Busse2009}{}%
Busse L, Wade AR, Carandini M (2009) Representation of concurrent stimuli by population activity in visual cortex. Neuron 64:931--942.

\leavevmode\hypertarget{ref-Campbell1965}{}%
Campbell FW, Green DG (1965) Monocular versus binocular visual acuity. Nature 208:191--192.

\leavevmode\hypertarget{ref-Carpenter2017}{}%
Carpenter B, Gelman A, Hoffman MD, Lee D, Goodrich B, Betancourt M, Brubaker MA, Guo J, Li P, Riddell A (2017) Stan: A probabilistic programming language. J Stat Softw 76.

\leavevmode\hypertarget{ref-Carroll2002}{}%
Carroll J, Neitz J, Neitz M (2002) Estimates of L:M cone ratio from ERG flicker photometry and genetics. Journal of Vision 2:1 Available at: \url{https://doi.org/10.1167/2.8.1} {[}Accessed June 5, 2023{]}.

\leavevmode\hypertarget{ref-CIE2006}{}%
CIE (2006) Fundamental chromaticity diagram with physiological axes. CIE Central Bureau. Available at: \url{https://cie.co.at/publications/fundamental-chromaticity-diagram-physiological-axes-part-1}.

\leavevmode\hypertarget{ref-CIE2018}{}%
CIE (2018) CIE system for metrology of optical radiation for ipRGC-influenced responses to light. CIE Central Bureau. Available at: \url{https://doi.org/10.25039/S026.2018}.

\leavevmode\hypertarget{ref-Dacey2003}{}%
Dacey DM, Peterson BB, Robinson FR, Gamlin PD (2003) Fireworks in the primate retina: In vitro photodynamics reveals diverse lgn-projecting ganglion cell types. Neuron 37:15--27 Available at: \url{https://www.sciencedirect.com/science/article/pii/S0896627302011431}.

\leavevmode\hypertarget{ref-Estevez1982}{}%
Estévez O, Spekreijse H (1982) The "silent substitution" method in visual research. Vision Research 22:681--691 Available at: \url{https://www.sciencedirect.com/science/article/pii/0042698982901043} {[}Accessed July 18, 2023{]}.

\leavevmode\hypertarget{ref-Gamlin2007}{}%
Gamlin PDR, McDougal DH, Pokorny J, Smith VC, Yau K-W, Dacey DM (2007) Human and macaque pupil responses driven by melanopsin-containing retinal ganglion cells. Vision Research 47:946--954 Available at: \url{https://www.sciencedirect.com/science/article/pii/S0042698906005682} {[}Accessed July 18, 2023{]}.

\leavevmode\hypertarget{ref-Hofer2005}{}%
Hofer H, Carroll J, Neitz J, Neitz M, Williams DR (2005) Organization of the human trichromatic cone mosaic. The Journal of Neuroscience 25:9669--9679 Available at: \url{https://doi.org/10.1523\%2Fjneurosci.2414-05.2005}.

\leavevmode\hypertarget{ref-Hubel1962}{}%
Hubel DH, Wiesel TN (1962) Receptive fields, binocular interaction and functional architecture in the cat's visual cortex. J Physiol 160:106--154.

\leavevmode\hypertarget{ref-Kaestner2019}{}%
Kaestner M, Maloney RT, Wailes-Newson KH, Bloj M, Harris JM, Morland AB, Wade AR (2019) Asymmetries between achromatic and chromatic extraction of 3D motion signals. Proc Natl Acad Sci U S A 116:13631--13640.

\leavevmode\hypertarget{ref-Kassner2014}{}%
Kassner M, Patera W, Bulling A (2014) Pupil: An open source platform for pervasive eye tracking and mobile gaze-based interaction. In: Proceedings of the 2014 ACM international joint conference on pervasive and ubiquitous computing: Adjunct publication. ACM. Available at: \url{https://doi.org/10.1145\%2F2638728.2641695}.

\leavevmode\hypertarget{ref-Kingdom2015}{}%
Kingdom FAA, Libenson L (2015) Dichoptic color saturation mixture: Binocular luminance contrast promotes perceptual averaging. J Vis 15:2.

\leavevmode\hypertarget{ref-Lucas2003}{}%
Lucas RJ, Hattar S, Takao M, Berson DM, Foster RG, Yau K-W (2003) Diminished pupillary light reflex at high irradiances in melanopsin-knockout mice. Science 299:245--247 Available at: \url{https://doi.org/10.1126\%2Fscience.1077293}.

\leavevmode\hypertarget{ref-Markwell2010}{}%
Markwell EL, Feigl B, Zele AJ (2010) Intrinsically photosensitive melanopsin retinal ganglion cell contributions to the pupillary light reflex and circadian rhythm. Clinical and Experimental Optometry 93:137--149 Available at: \url{https://doi.org/10.1111\%2Fj.1444-0938.2010.00479.x}.

\leavevmode\hypertarget{ref-Martin2023}{}%
Martin JT, Boynton GM, Baker DH, Wade AR, Spitschan M (2023) PySilSub: An open-source python toolbox for implementing the method of silent substitution in vision and nonvisual photoreception research. Journal of Vision 23:10 Available at: \url{https://doi.org/10.1167\%2Fjov.23.7.10}.

\leavevmode\hypertarget{ref-Martin2022}{}%
Martin JT, Pinto J, Bulte D, Spitschan M (2022) PyPlr: A versatile, integrated system of hardware and software for researching the human pupillary light reflex. Behavior Research Methods 54:2720--2739.

\leavevmode\hypertarget{ref-Mathot2018}{}%
Mathôt S (2018) Pupillometry: Psychology, physiology, and function. J Cogn 1:16.

\leavevmode\hypertarget{ref-McDougal2010}{}%
McDougal DH, Gamlin PD (2010) The influence of intrinsically-photosensitive retinal ganglion cells on the spectral sensitivity and response dynamics of the human pupillary light reflex. Vision Research 50:72--87 Available at: \url{https://www.sciencedirect.com/science/article/pii/S0042698909004763} {[}Accessed July 18, 2023{]}.

>>>>

Different classes of retinal photoreceptors, including cones, rods and melanopsin-containing intrinsically photosensitive retinal ganglion cells (ipRGCs), are directly involved in controlling and maintaining the size of the pupils (e.g.~McDougal and Gamlin, \protect\hyperlink{ref-McDougal2010}{2010}; Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Murray et al., \protect\hyperlink{ref-Murray2018}{2018}; Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}; Woelders et al., \protect\hyperlink{ref-Woelders2018}{2018}). Cones drive the initial rapid constriction of the pupils (Mathôt, \protect\hyperlink{ref-Mathot2018}{2018}), while the slower and longer activation of the ipRGCs maintains constriction over a prolonged period of time and regulates the post-illumination pupillary response (Markwell et al., \protect\hyperlink{ref-Markwell2010}{2010}; McDougal and Gamlin, \protect\hyperlink{ref-McDougal2010}{2010}). The ipRGCs are a recently discovered photoreceptor class (Provencio et al., \protect\hyperlink{ref-Provencio2000}{2000}) that express the photopigment melanopsin, and are involved in the regulation of the circadian rhythm (Panda et al., \protect\hyperlink{ref-Panda2002}{2002}; Ruby et al., \protect\hyperlink{ref-Ruby2002}{2002}), forming a major input to the PON (Dacey et al., \protect\hyperlink{ref-Dacey2003}{2003}). The first direct evidence of the involvement of the ipRGCs in the PLR was shown in melanopsin knockout rats by Lucas et al. (\protect\hyperlink{ref-Lucas2003}{2003}), resulting in the loss of the intrinsic photosensitivity of the cells and a reduced pupil constriction. Similar behaviour was later observed in primates and humans (Gamlin et al., \protect\hyperlink{ref-Gamlin2007}{2007}) using silent substitution (see Methods), where it was demonstrated that the PLR continues during light presentation even when cone and rod signalling is blocked, indicating the primary role of the ipRGCs in maintaining pupil constriction over a prolonged time.

Binocular combination has been extensively studied in visual perception, where it is mediated by neurons in primary visual cortex (Hubel and Wiesel, \protect\hyperlink{ref-Hubel1962}{1962}). For pattern vision, binocular summation occurs at threshold, such that a lower contrast is required to detect a target shown to both eyes than a target shown to one eye (Baker et al., \protect\hyperlink{ref-Baker2018}{2018}; Campbell and Green, \protect\hyperlink{ref-Campbell1965}{1965}). At higher contrasts, the phenomenon of `ocularity invariance' is observed, in which the response to monocularly- and binocularly-presented patterns is equal (Baker et al., \protect\hyperlink{ref-Baker2007}{2007}; Moradi and Heeger, \protect\hyperlink{ref-Moradi2009}{2009}). This is explained by a process of interocular suppression that cancels out the additional excitatory drive caused by stimulating two eyes. Our recent work (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}) showed that cortical signal combination for luminance flicker is substantially more linear than for spatial patterns, whereas there is evidence of interocular suppression in the pupil pathway. Here we measure the amplitude of pupil modulations in response to flickering stimuli presented as light flux, or directed towards the L-M cone, S-cone, and melanopsin pathways (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and in a dichoptic masking condition where the two eyes are stimulated at different frequencies. The results are interpreted using a contemporary model of binocular vision (Meese et al., \protect\hyperlink{ref-Meese2006}{2006}) implemented within a hierarchical Bayesian framework. These results previously formed part of a doctoral thesis by the first author (Segala, \protect\hyperlink{ref-SegalaThesis2023}{2023}).

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies (e.g., Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Murray et al., \protect\hyperlink{ref-Murray2018}{2018}; Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}, \protect\hyperlink{ref-Spitschan2014}{2014}; Zele et al., \protect\hyperlink{ref-Zele2018}{2018}). The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study (Segala et al., \protect\hyperlink{ref-Segala2023}{2023}): monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean light flux. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany, Kassner et al., \protect\hyperlink{ref-Kassner2014}{2014}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

Silent substitution stimuli were prepared and calibrated for each participant with custom Python software (Martin et al., \protect\hyperlink{ref-Martin2023}{2023}) and Python scripts. Estimates of photoreceptor spectral sensitivities for each participant were constructed from the known photopigment absorbance spectra (Stockman and Sharpe, \protect\hyperlink{ref-Stockman2000}{2000}), taking account of the peak axial density of the respective photopigments, as well as lens (Pokorny et al., \protect\hyperlink{ref-Pokorny1987}{1987}; Stockman and Sharpe, \protect\hyperlink{ref-Stockman2000}{2000}; Stockman et al., \protect\hyperlink{ref-Stockman1999}{1999}) and macular pigment density (Bone et al., \protect\hyperlink{ref-Bone1988}{1988}; Stockman et al., \protect\hyperlink{ref-Stockman1999}{1999}), in accordance with the field size and age-dependant CIEPO06 observer model (CIE, \protect\hyperlink{ref-CIE2006}{2006}). The melanopic and rhodopic action spectra of the 32-year-old standard observer were taken from CIE S 026 (CIE, \protect\hyperlink{ref-CIE2018}{2018}) and then adjusted for age-related lens transmittance with a spectral correction function, in line with the standard. Macular pigment correction was not applied to the rhodopic and melanopic action spectra because rods are not present at the fovea and ipRGCs sit above the retinal pigment layer (Trieschmann et al., \protect\hyperlink{ref-Trieschmann2007}{2007}).

In the light flux experiment, the stimulus intensity was increased and decreased relative to the background, which we expected to modulate all photoreceptor classes (see Figure \ref{fig:spectraplots}e,i). In the L-M cone modulation experiment, we used silent substitution to increase the L-cone activity, and simultaneously decrease the M-cone activity, during the first half-cycle of the sine wave. In the second half-cycle the polarity of the modulation reversed (see Figure \ref{fig:spectraplots}f,j). The maximum available L-M contrast was approximately 10\%. In the S-cone modulation experiment, we increased and decreased S-cone-directed signals, whilst keeping activity in the other photoreceptors constant (see Figure \ref{fig:spectraplots}g,k). Our system allowed a maximum contrast of 45\%. Finally, in the melanopsin experiment, we modulated the activity of the melanopsin-containing intrinsically photoreceptive retinal ganglion cells, whilst keeping cone activity constant (see Figure \ref{fig:spectraplots}h,l). The maximum available melanopsin contrast was 22\%. We assume that the activity of rods was constant at the high background luminance intensity used here, and so did not attempt to silence rod activity in any condition, as this would have greatly reduced the available dynamic range. Splatter on nominally silenced photoreceptors was very small (see Figure \ref{fig:spectraplots}j-l), well below the levels that would be expected to generate measurable pupil modulations, although we can observe that the rods may not be saturated in the L-M and melanopsin-directed stimuli (Figure \ref{fig:spectraplots}j and \ref{fig:spectraplots}l) and could intrude in these two experiments. We also estimated activation of penumbral L and M cones (Barrionuevo and Cao, \protect\hyperlink{ref-Barrionuevo2016}{2016}; Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}; Zele et al., \protect\hyperlink{ref-Zele2019}{2019}), which was minimal (\(\le 1.5%
\) contrast; less than splatter on the open-field cones) for melanopsin-directed stimuli. We note that the temporal frequency of our modulation (0.5Hz) is well below the range where penumbral cone activation can elicit visible percepts, and that such percepts fade after around 1 second (Spitschan et al., \protect\hyperlink{ref-Spitschan2015}{2015}), and do not typically affect pupil responses (Spitschan et al., \protect\hyperlink{ref-Spitschan2014}{2014}).

@@

Before the start of each experiment, participants adjusted the objective planes of the lenses with the help of the experimenter until the stimulus was in focus and they perceived the two pieces of blackout material as one fused disc. Pupil responses to binocular temporal contrast modulations were examined in a factorial design that combined six ocular conditions and five temporal contrast levels: 6, 12, 24, 48 and 96\% of the available dynamic range. This design, similar to that used in our previous studies (Baker et al., \protect\hyperlink{ref-Baker2020}{2020}; Baker and Wade, \protect\hyperlink{ref-Baker2017}{2017}; Segala et al., \protect\hyperlink{ref-Segala2023}{2023}), was applied in four separate experiments, each with a different mode of photoreceptor stimulation. In the first three conditions, the discs flickered at 0.5 Hz, in either a monocular, binocular or dichoptic arrangement. In the dichoptic condition the non-target eye saw a flickering fixed contrast of 48\% of the available dynamic range. In the remaining three conditions (the cross-frequency conditions) one eye's disc flickered at 0.4 Hz, and the other eye's disc flickered at 0.5 Hz. This included monocular responses at 0.4 Hz, as well as binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 0.5 Hz, mask contrast of 48\% at 0.4 Hz in the other eye) arrangements. We counterbalanced presentation of the target stimulus across the left and right eyes.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Baker2021}{}%
Baker DH. 2021. Statistical analysis of periodic data in neuroscience. \emph{Neurons, Behavior, Data analysis, and Theory} \textbf{5}. doi:\href{https://doi.org/10.51628/001c.27680}{10.51628/001c.27680}

\leavevmode\hypertarget{ref-Baker2024}{}%
Baker DH, Hansford KJ, Segala FG, Morsi AY, Huxley RJ, Martin JT, Rockman M, Wade AR. 2024. Binocular integration of chromatic and luminance signals. \emph{J Vis} \textbf{24(12)}:7. doi:\href{https://doi.org/10.1167/jov.24.12.7}{10.1167/jov.24.12.7}

\leavevmode\hypertarget{ref-Baker2018}{}%
Baker DH, Lygo FA, Meese TS, Georgeson MA. 2018. Binocular summation revisited: Beyond \(\sqrt{2}\). \emph{Psychol Bull} \textbf{144}:1186--1199. doi:\href{https://doi.org/10.1037/bul0000163}{10.1037/bul0000163}

\leavevmode\hypertarget{ref-Baker2007}{}%
Baker DH, Meese TS, Georgeson MA. 2007. Binocular interaction: Contrast matching and contrast discrimination are predicted by the same model. \emph{Spat Vis} \textbf{20}:397--413. doi:\href{https://doi.org/10.1163/156856807781503622}{10.1163/156856807781503622}

\leavevmode\hypertarget{ref-Baker2020}{}%
Baker DH, Vilidaite G, McClarnon E, Valkova E, Bruno A, Millman RE. 2020. Binaural summation of amplitude modulation involves weak interaural suppression. \emph{Sci Rep} \textbf{10}:3560. doi:\href{https://doi.org/10.1038/s41598-020-60602-5}{10.1038/s41598-020-60602-5}

\leavevmode\hypertarget{ref-Baker2017}{}%
Baker DH, Wade AR. 2017. Evidence for an optimal algorithm underlying signal combination in human visual cortex. \emph{Cereb Cortex} \textbf{27}:254--264. doi:\href{https://doi.org/10.1093/cercor/bhw395}{10.1093/cercor/bhw395}

\leavevmode\hypertarget{ref-Barrionuevo2016}{}%
Barrionuevo PA, Cao D. 2016. Luminance and chromatic signals interact differently with melanopsin activation to control the pupil light response. \emph{Journal of Vision} \textbf{16(11)}:29. doi:\href{https://doi.org/10.1167/16.11.29}{10.1167/16.11.29}

\leavevmode\hypertarget{ref-Bone1988}{}%
Bone RA, Landrum JT, Fernandez L, Tarsis SL. 1988. Analysis of the macular pigment by HPLC: Retinal distribution and age study. \emph{Investigative Ophthalmology \& Visual Science} \textbf{29}:843--849.

\leavevmode\hypertarget{ref-Busse2009}{}%
Busse L, Wade AR, Carandini M. 2009. Representation of concurrent stimuli by population activity in visual cortex. \emph{Neuron} \textbf{64}:931--42. doi:\href{https://doi.org/10.1016/j.neuron.2009.11.004}{10.1016/j.neuron.2009.11.004}

\leavevmode\hypertarget{ref-Campbell1965}{}%
Campbell FW, Green DG. 1965. Monocular versus binocular visual acuity. \emph{Nature} \textbf{208}:191--2. doi:\href{https://doi.org/10.1038/208191a0}{10.1038/208191a0}

\leavevmode\hypertarget{ref-Carpenter2017}{}%
Carpenter B, Gelman A, Hoffman MD, Lee D, Goodrich B, Betancourt M, Brubaker MA, Guo J, Li P, Riddell A. 2017. Stan: A probabilistic programming language. \emph{J Stat Softw} \textbf{76}. doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}

\leavevmode\hypertarget{ref-Carroll2002}{}%
Carroll J, Neitz J, Neitz M. 2002. Estimates of L:M cone ratio from ERG flicker photometry and genetics. \emph{Journal of Vision} \textbf{2}:1. doi:\href{https://doi.org/10.1167/2.8.1}{10.1167/2.8.1}

\leavevmode\hypertarget{ref-CIE2018}{}%
CIE. 2018. CIE system for metrology of optical radiation for ipRGC-influenced responses to light. CIE Central Bureau. doi:\href{https://doi.org/10.25039/S026.2018}{10.25039/S026.2018}

\leavevmode\hypertarget{ref-CIE2006}{}%
CIE. 2006. Fundamental chromaticity diagram with physiological axes. CIE Central Bureau.

\leavevmode\hypertarget{ref-Dacey2003}{}%
Dacey DM, Peterson BB, Robinson FR, Gamlin PD. 2003. Fireworks in the primate retina: In vitro photodynamics reveals diverse lgn-projecting ganglion cell types. \emph{Neuron} \textbf{37}:15--27. doi:\href{https://doi.org/https://doi.org/10.1016/S0896-6273(02)01143-1}{https://doi.org/10.1016/S0896-6273(02)01143-1}

\leavevmode\hypertarget{ref-Estevez1982}{}%
Estévez O, Spekreijse H. 1982. The "silent substitution" method in visual research. \emph{Vision Research} \textbf{22}:681--691. doi:\href{https://doi.org/10.1016/0042-6989(82)90104-3}{10.1016/0042-6989(82)90104-3}

\leavevmode\hypertarget{ref-Gamlin2007}{}%
Gamlin PDR, McDougal DH, Pokorny J, Smith VC, Yau K-W, Dacey DM. 2007. Human and macaque pupil responses driven by melanopsin-containing retinal ganglion cells. \emph{Vision Research} \textbf{47}:946--954. doi:\href{https://doi.org/10.1016/j.visres.2006.12.015}{10.1016/j.visres.2006.12.015}

\leavevmode\hypertarget{ref-Hofer2005}{}%
Hofer H, Carroll J, Neitz J, Neitz M, Williams DR. 2005. Organization of the human trichromatic cone mosaic. \emph{The Journal of Neuroscience} \textbf{25}:9669--9679. doi:\href{https://doi.org/10.1523/jneurosci.2414-05.2005}{10.1523/jneurosci.2414-05.2005}

\leavevmode\hypertarget{ref-Hubel1962}{}%
Hubel DH, Wiesel TN. 1962. Receptive fields, binocular interaction and functional architecture in the cat's visual cortex. \emph{J Physiol} \textbf{160}:106--54. doi:\href{https://doi.org/10.1113/jphysiol.1962.sp006837}{10.1113/jphysiol.1962.sp006837}

\leavevmode\hypertarget{ref-Kaestner2019}{}%
Kaestner M, Maloney RT, Wailes-Newson KH, Bloj M, Harris JM, Morland AB, Wade AR. 2019. Asymmetries between achromatic and chromatic extraction of 3D motion signals. \emph{Proc Natl Acad Sci U S A} \textbf{116}:13631--13640. doi:\href{https://doi.org/10.1073/pnas.1817202116}{10.1073/pnas.1817202116}

\leavevmode\hypertarget{ref-Kassner2014}{}%
Kassner M, Patera W, Bulling A. 2014. Pupil: An open source platform for pervasive eye tracking and mobile gaze-based interaction.Proceedings of the 2014 ACM International Joint Conference on Pervasive and Ubiquitous Computing: Adjunct Publication. ACM. doi:\href{https://doi.org/10.1145/2638728.2641695}{10.1145/2638728.2641695}

\leavevmode\hypertarget{ref-Kingdom2015}{}%
Kingdom FAA, Libenson L. 2015. Dichoptic color saturation mixture: Binocular luminance contrast promotes perceptual averaging. \emph{J Vis} \textbf{15}:2. doi:\href{https://doi.org/10.1167/15.5.2}{10.1167/15.5.2}

\leavevmode\hypertarget{ref-Lucas2003}{}%
Lucas RJ, Hattar S, Takao M, Berson DM, Foster RG, Yau K-W. 2003. Diminished pupillary light reflex at high irradiances in melanopsin-knockout mice. \emph{Science} \textbf{299}:245--247. doi:\href{https://doi.org/10.1126/science.1077293}{10.1126/science.1077293}
